---
title: "Load & Review Data v2 (get all feet sorted)"
author: "Wesley"
date: "2022/03/28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load all of the Foot Data

To start, we need to load all of the Excel files. To do this, we will use the
readxl package. In addition, we're going to need to manipulate strings in 
the following, so let's pre-emptively load that package.

```{r libraries, error = FALSE, message = FALSE, warning = FALSE}
library("readxl")
library("stringr")
```

### Start with Foot 1, as Template

Foot 1 has many sheets, each of which is a nicely organized listing of VOCs
as detected by the GCxGC system's built-in scanning software, and exported. 
We will load them all. 

```{r load_data}
sheets <- excel_sheets("./dat/foot1.xlsx")  # note: spaces in names
foot1 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "./dat/foot1.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(foot1) <- sheets
```

Having things organized by sample is a good thing, but we really will
find it annoying finding things long-term. In the following, I will take each
of the lists of compounds relevant to Foot 1, and merge them into a single
data.frame object, so we can then filter that frame down to get a specific
compound, or a specific day/observation, and so on.

Before we do this, the first sheets has the observations duplicated across
the columns (wups!). Removing that:

```{r remove_sheet1}
foot1[[1]] <- foot1[[1]][, 1:ncol(foot1[[2]])]
names(foot1[[1]]) <- names(foot1[[2]])  # fix names of sheet 1
```

```{r merge_sheets}
foot1m <- do.call("rbind", foot1)
names(foot1m) <- c("PeakNum", "Sample", "Name", "Formula", "RTs", "Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN",
                   "Type", "Conc", "ConcUnits")
```

A quick print of the data structure shows that the RTs, QuantMass and BaseMass
won't be usable as numbers. The RTs need to be split, as well.

```{r split_RTs}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(foot1m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(foot1m$RTs, ","), FUN = "[[", 2)
         )
       )
foot1m <- data.frame(foot1m[, c(1:4, 6:15)], RT1 = RT1, RT2 = RT2)
```

```{r samples}
unique(foot1m$Sample)
save(file = "./dat/foot1.rda", foot1m)
```

### Now Do Foot 2

```{r load_data2}
sheets <- excel_sheets("./dat/foot2.xlsx")  # note: spaces in names
foot2 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "./dat/foot2.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(foot2) <- sheets

foot2m <- do.call("rbind", foot2)
names(foot2m) <- c("PeakNum", "Sample", "Name", "Formula", "RTs", "Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN",
                   "Type", "Conc", "ConcUnits")
```

```{r split_RTs2}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(foot2m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(foot2m$RTs, ","), FUN = "[[", 2)
         )
       )
foot2m <- data.frame(foot2m[, c(1:4, 6:15)], RT1 = RT1, RT2 = RT2)
```

```{r samples2}
unique(foot2m$Sample)
save(file = "./dat/foot2.rda", foot2m)
```

### Now Do Foot 3

```{r load_data3}
sheets <- excel_sheets("./dat/foot3.xlsx")  # note: spaces in names
foot3 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "./dat/foot3.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(foot3) <- sheets

foot3m <- do.call("rbind", foot3)
names(foot3m) <- c("PeakNum", "Sample", "Name", "Formula", "RTs", "Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN",
                   "Type", "Conc", "ConcUnits")
```

```{r split_RTs3}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(foot3m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(foot3m$RTs, ","), FUN = "[[", 2)
         )
       )
foot3m <- data.frame(foot3m[, c(1:4, 6:15)], RT1 = RT1, RT2 = RT2)
```

```{r samples2}
unique(foot3m$Sample)
save(file = "./dat/foot3.rda", foot3m)
```

### Now Do Foot 4

```{r load_data4}
sheets <- excel_sheets("./dat/foot4.xlsx")  # note: spaces in names
foot4 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "./dat/foot4.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(foot4) <- sheets

foot4m <- do.call("rbind", foot4)
names(foot4m) <- c("PeakNum", "Sample", "Name", "Formula", "RTs", "Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN",
                   "Type", "Conc", "ConcUnits")
```

```{r split_RTs4}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(foot4m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(foot4m$RTs, ","), FUN = "[[", 2)
         )
       )
foot4m <- data.frame(foot4m[, c(1:4, 6:15)], RT1 = RT1, RT2 = RT2)
```

```{r samples2}
unique(foot4m$Sample)
save(file = "./dat/foot4.rda", foot4m)
```

### Now Do Foot 5

```{r load_data5}
sheets <- excel_sheets("./dat/foot5.xlsx")  # note: spaces in names
foot5 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "./dat/foot5.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(foot5) <- sheets

foot5m <- do.call("rbind", foot5)
names(foot5m) <- c("PeakNum", "Sample", "Name", "Formula", "RTs", "Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN",
                   "Type", "Conc", "ConcUnits")
```

```{r split_RTs4}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(foot5m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(foot5m$RTs, ","), FUN = "[[", 2)
         )
       )
foot5m <- data.frame(foot5m[, c(1:4, 6:15)], RT1 = RT1, RT2 = RT2)
```

```{r samples2}
unique(foot5m$Sample)
save(file = "./dat/foot5.rda", foot5m)
```

### Now Do Foot 6

```{r load_data6}
sheets <- excel_sheets("./dat/foot6.xlsx")  # note: spaces in names
foot6 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "./dat/foot6.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(foot6) <- sheets

foot6m <- do.call("rbind", foot6)
names(foot6m) <- c("PeakNum", "Sample", "Name", "Formula", "RTs", "Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN",
                   "Type", "Conc", "ConcUnits")
```

```{r split_RTs4}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(foot6m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(foot6m$RTs, ","), FUN = "[[", 2)
         )
       )
foot6m <- data.frame(foot6m[, c(1:4, 6:15)], RT1 = RT1, RT2 = RT2)
```

```{r samples2}
unique(foot6m$Sample)
save(file = "./dat/foot6.rda", foot6m)
```

### Now Do Foot 7

```{r load_data7}
sheets <- excel_sheets("./dat/foot7.xlsx")  # note: spaces in names
foot7 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "./dat/foot7.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(foot7) <- sheets

foot7m <- do.call("rbind", foot7)
names(foot7m) <- c("PeakNum", "Sample", "Name", "Formula", "RTs", "Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN",
                   "Type", "Conc", "ConcUnits")
```

```{r split_RTs4}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(foot7m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(foot7m$RTs, ","), FUN = "[[", 2)
         )
       )
foot7m <- data.frame(foot7m[, c(1:4, 6:15)], RT1 = RT1, RT2 = RT2)
```

```{r samples2}
unique(foot7m$Sample)
save(file = "./dat/foot7.rda", foot7m)
```

### Now Do Foot 8

```{r load_data8}
sheets <- excel_sheets("./dat/foot8.xlsx")  # note: spaces in names
foot8 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "./dat/foot8.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(foot8) <- sheets

foot8m <- do.call("rbind", foot8)
names(foot8m) <- c("PeakNum", "Sample", "Name", "Formula", "RTs", "Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN",
                   "Type", "Conc", "ConcUnits")
```

```{r split_RTs4}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(foot8m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(foot8m$RTs, ","), FUN = "[[", 2)
         )
       )
foot8m <- data.frame(foot8m[, c(1:4, 6:15)], RT1 = RT1, RT2 = RT2)
```

```{r samples2}
unique(foot8m$Sample)
save(file = "./dat/foot8.rda", foot8m)
```

### Now Do Foot 9

```{r load_data9}
sheets <- excel_sheets("./dat/foot9.xlsx")  # note: spaces in names
foot9 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "./dat/foot9.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(foot9) <- sheets

foot9m <- do.call("rbind", foot9)
names(foot9m) <- c("PeakNum", "Sample", "Name", "Formula", "RTs", "Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN",
                   "Type", "Conc", "ConcUnits")
```

```{r split_RTs4}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(foot9m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(foot9m$RTs, ","), FUN = "[[", 2)
         )
       )
foot9m <- data.frame(foot9m[, c(1:4, 6:15)], RT1 = RT1, RT2 = RT2)
```

```{r samples2}
unique(foot9m$Sample)
save(file = "./dat/foot9.rda", foot9m)
```

