---
title: "General Function Cleaning"
author: "Wesley Burr"
date: "28/03/2022"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(stringr)
```

```{r function}
##
#  clean_common: Function to select, filter, clean, filter and merge 
#  compounds using the logic of:
#  * unique to the samples, not the control, for ALL samples; or
#  * in both samples and control, but much stronger in the control
# 
#  Inputs:
#  * dat: data.frame sourced from merging spreadsheets of GCxGC output
#  * sample_names: names of specific samples (e.g., SS5_Foot_1_a)
#  * control_names: names of specific control samples (e.g., SS5_Foot_1_Control_a)
#  * ratio_par: cut-off for the logic of "in both samples and control" - if this
#      is set very large, will eliminate cross-overs.
#
#  Returns:
#  * samples_keep: list of full data.frames for individual replicates, cleaned
#      down to relevant compounds using above logic
##
clean_common <- function(dat,
                         sample_names,
                         control_names,
                         ratio_par = 2.0) {
  
  samples <- vector("list", length = length(sample_names))
  names(samples) <- sample_names
  controls <- vector("list", length = length(control_names))
  names(controls) <- control_names
  
  # Extract specific samples and controls of interest and
  # dump all but the largest Area example of each compound
  for(j in 1:length(sample_names)) {
    samples[[j]] <- dat %>% subset(Sample == sample_names[j]) %>%
                      group_by(Name) %>%
                      filter(Area == max(Area)) %>%
                      ungroup() %>% filter(substr(Name, 1, 4) != "Peak")
    samples[[j]] <- samples[[j]][!duplicated(samples[[j]]$Name), ]
  } 
  for(j in 1:length(control_names)) {
    controls[[j]] <- dat %>% subset(Sample == control_names[j]) %>%
                      group_by(Name) %>%
                      filter(Area == max(Area)) %>%
                      ungroup() %>% filter(substr(Name, 1, 4) != "Peak")
    controls[[j]] <- controls[[j]][!duplicated(controls[[j]]$Name), ]
  } 
  # merge controls
  control <- do.call("rbind", controls)
  control <- control %>% group_by(Name) %>%
                      filter(Area == max(Area)) %>%
                      ungroup() 
  control <- control[!duplicated(control$Name), ]

  # Find compounds that are in each sample that are also in control
  samples_keep <- samples
  for(j in 1:length(sample_names)) {
    samp <- samples[[j]] %>% filter(samples[[j]]$Name %in% control$Name)
    cont <- control %>% filter(control$Name %in% samples[[j]]$Name)
   
    # ratio is high enough to keep 
    samp_SN <- unlist(samp[order(samp$Name), "PeakSN"])
    cont_SN <- unlist(cont[order(cont$Name), "PeakSN"])
    contrib1 <- samp %>% filter((samp_SN / cont_SN) > ratio_par)
   
    # also, compounds that are *not* in the controls 
    contrib2 <- samples[[j]] %>% filter(!(samples[[j]]$Name %in% control$Name))
    samples_keep[[j]] <- rbind(contrib1, contrib2)
  }
  names(samples_keep) <- sample_names
  samples_keep
}


##
#
#  join_common: Function which takes output of clean_common above,
#    and merges based on common presence across all replicates of compounds. 
#
#  Inputs:
#  * compounds: list of data.frames, 16 columns as in the spreadsheets
#  
#  Outputs:
#  * common: merged, simplified data.frame, created via inner_join of data.frames after filtering.
##
join_common <- function(compounds) {
  n_samp <- length(compounds)
  subset_compounds <- vector("list", length = n_samp)
  for(j in 1:n_samp) {
    subset_compounds[[j]] <- compounds[[j]]
    if(n_samp > 1) {
      for(k in (1:n_samp)[-j]) {
        subset_compounds[[j]] <- subset_compounds[[j]] %>%
                                   subset(subset_compounds[[j]]$Name %in% compounds[[k]]$Name)
      }
    }
    subset_compounds[[j]] <- subset_compounds[[j]] %>% select(Name, Area, PeakSN)
  }
  
  # Join first two, if they exist
  if(n_samp > 1) {
    common <- inner_join(x = subset_compounds[[1]], y = subset_compounds[[2]], by = "Name")
    if(n_samp >= 3) {
      for(j in 3:n_samp) {
        common <- inner_join(x = common, y = subset_compounds[[j]], by = "Name")  
      }
    }
  } else {
    common <- subset_compounds[[1]][, c("Name", "Area", "PeakSN")]
  }
  names(common) <- c("Name", paste0(c("Area_", "PeakSN_"), rep(1:n_samp, each = 2)))
  common
}
```

Let's test this:

```{r}
load("./dat/foot1.rda")
foot1m$Sample <- str_replace_all(foot1m$Sample, " ", "_")

testing <- clean_common(foot1m,
                sample_names = c("SS2_Foot_1_a", "SS2_Foot_1_b"),
                control_names = c("SS2_Foot_1_Control"),
                ratio_par = 2.0)
common <- join_common(testing)
str(common)
```

### Testing a Triplet

Testing to make sure this method works ok for triplets (e.g., SS5, where
there are 3 replicate samples and 3 controls). It does! The above code should be 
sufficient to clean up things tremendously: two function calls, and we have
the common compounds across $N$ replicates and $M$ controls, using the same logic
throughout. Very easy. 

```{r}
testing <- clean_common(foot1m,
                sample_names = c("SS5_Foot_1_a", "SS5_Foot_1_b", "SS5_Foot_1_c"),
                control_names = c("SS5_Foot_1_Control_a", "SS5_Foot_1_Control_b", "SS5_Foot_1_Control_c"),
                ratio_par = 2.0)
common <- join_common(testing)
str(common)
```

# Working Through All the Feet

We've already done Foot 1, 2 and 5, but we'll re-do them here so it's for-sure
consistent.

```{r foot1}
load("./dat/foot1.rda")
foot1m$Sample <- str_replace_all(foot1m$Sample, " ", "_")

foot1_SS2 <- join_common( clean_common(foot1m,
                sample_names = c("SS2_Foot_1_a", "SS2_Foot_1_b"),
                control_names = c("SS2_Foot_1_Control"),
                ratio_par = 2.0) )
foot1_SS3 <- join_common( clean_common(foot1m,
                sample_names = c("SS3_Foot_1_a", "SS3_Foot_1_b", "SS3_Foot_1_c"),
                control_names = c("SS3_Foot_1_Control"),
                ratio_par = 2.0) )
foot1_SS4 <- join_common( clean_common(foot1m,
                sample_names = c("SS4_Foot_1_a", "SS4_Foot_1_b", "SS4_Foot_1_c"),
                control_names = c("SS4_Foot_1_Control_a", "SS4_Foot_1_Control_b", "SS4_Foot_1_Control_c"),
                ratio_par = 2.0) )
foot1_SS5 <- join_common( clean_common(foot1m,
                sample_names = c("SS5_Foot_1_a", "SS5_Foot_1_b", "SS5_Foot_1_c"),
                control_names = c("SS5_Foot_1_Control_a", "SS5_Foot_1_Control_b", "SS5_Foot_1_Control_c"),
                ratio_par = 2.0) )
foot1_SS6 <- join_common( clean_common(foot1m,
                sample_names = c("SS6_Foot_1_a", "SS6_Foot_1_b"),
                control_names = c("SS6_Foot_1_Control_a", "SS6_Foot_1_Control_b"),
                ratio_par = 2.0) )
```


```{r foot2}
load("./dat/foot2.rda")
foot2m$Sample <- str_replace_all(foot2m$Sample, " ", "_")

foot2_SS2 <- join_common( clean_common(foot2m,
                sample_names = c("SS2_Foot_2_a", "SS2_Foot_2_b"),
                control_names = c("SS2_Foot_2_Control"),
                ratio_par = 2.0) )
foot2_SS3 <- join_common( clean_common(foot2m,
                sample_names = c("SS3_Foot_2_a", "SS3_Foot_2_b", "SS3_Foot_2_c"),
                control_names = c("SS3_Foot_2_Control"),
                ratio_par = 2.0) )
foot2_SS4 <- join_common( clean_common(foot2m,
                sample_names = c("SS4_Foot_2_a", "SS4_Foot_2_b", "SS4_Foot_2_c"),
                control_names = c("SS4_Foot_2_Control_a", "SS4_Foot_2_Control_b", "SS4_Foot_2_Control_c"),
                ratio_par = 2.0) )
foot2_SS5 <- join_common( clean_common(foot2m,
                sample_names = c("SS5_Foot_2_a", "SS5_Foot_2_b", "SS5_Foot_2_c"),
                control_names = c("SS5_Foot_2_Control_a", "SS5_Foot_2_Control_b", "SS5_Foot_2_Control_c"),
                ratio_par = 2.0) )
foot2_SS6 <- join_common( clean_common(foot2m,
                sample_names = c("SS6_Foot_2_a", "SS6_Foot_2_b"),
                control_names = c("SS6_Foot_2_Control_a", "SS6_Foot_2_Control_b"),
                ratio_par = 2.0) )
```

```{r foot3}
load("./dat/foot3.rda")
foot3m$Sample <- str_replace_all(foot3m$Sample, " ", "_")

foot3_SS5 <- join_common( clean_common(foot3m,
                sample_names = c("SS5_Bone_3_a", "SS5_Foot_3_b"),
                control_names = c("SS5_Foot_3_Control_a_(PVC_+_big_jar_a)",
                                  "SS5_Foot_3_Control_b_(PVC_+_zip_lock_a)"),
                ratio_par = 2.0) )
foot3_SS6 <- join_common( clean_common(foot3m,
                sample_names = c("SS6_Foot_3_a", "SS6_Foot_3_b", "SS6_Foot_3_c"),
                control_names = c("SS6_Foot_3_Control_a_(PVC_+_big_glass_jar_a)",
                                  "SS6_Foot_3_Control_a_(PVC_+_zip_lock_a)"),
                ratio_par = 2.0) )
```

```{r foot4}
load("./dat/foot4.rda")
foot4m$Sample <- str_replace_all(foot4m$Sample, " ", "_")

foot4_SS5 <- join_common( clean_common(foot4m,
                sample_names = c("SS5_Foot_4_a", "SS5_Foot_4_b"),
                control_names = c("SS5_Foot_4_Control_a_(PVC_+_big_jar_a)",
                                  "SS5_Foot_4_Control_b_(PVC_+_zip_lock_a)"),
                ratio_par = 2.0) )
foot4_SS6 <- join_common( clean_common(foot4m,
                sample_names = c("SS6_Foot_4_a", "SS6_Foot_4_b", "SS6_Foot_4_c"),
                control_names = c("SS6_Foot_4_Control_a_(PVC_+_big_glass_jar_a)",
                                  "SS6_Foot_4_Control_a_(PVC_+_zip_lock_a)"),
                ratio_par = 2.0) )
```

```{r foot5}
load("./dat/foot5.rda")
foot5m$Sample <- str_replace_all(foot5m$Sample, " ", "_")

foot5_SS2 <- join_common( clean_common(foot5m,
                sample_names = c("SS2_Foot_5_a"),
                control_names = c("SS2_Foot_5_Control"),
                ratio_par = 2.0) )
foot5_SS3 <- join_common( clean_common(foot5m,
                sample_names = c("SS3_Foot_5_a", "SS3_Foot_5_b", "SS3_Foot_5_c"),
                control_names = c("SS3_Foot_5_Control"),
                ratio_par = 2.0) )
foot5_SS4 <- join_common( clean_common(foot5m,
                sample_names = c("SS4_Foot_5_a", "SS4_Foot_5_b", "SS4_Foot_5_c"),
                control_names = c("SS4_Foot_5_Control_a", "SS4_Foot_5_Control_b", 
                                  "SS4_Foot_5_Control_c"),
                ratio_par = 2.0) )
foot5_SS5 <- join_common( clean_common(foot5m,
                sample_names = c("SS5_Foot_5_a", "SS5_Foot_5_b", "SS5_Foot_5_c"),
                control_names = c("SS5_Foot_5_Control_a", "SS5_Foot_5_Control_b", 
                                  "SS5_Foot_5_Control_c"),
                ratio_par = 2.0) )
foot5_SS6 <- join_common( clean_common(foot5m,
                sample_names = c("SS6_Foot_5_a", "SS6_Foot_5_b"),
                control_names = c("SS6_Foot_5_Control_a", "SS6_Foot_5_Control_b", 
                                  "SS6_Foot_5_Control_c"),
                ratio_par = 2.0) )
```

```{r foot6}
load("./dat/foot6.rda")
foot6m$Sample <- str_replace_all(foot6m$Sample, " ", "_")

foot6_SS5 <- join_common( clean_common(foot6m,
                sample_names = c("SS5_Foot_6_a", "SS5_Foot_6_b", "SS5_Foot_6_c"),
                control_names = c("SS5_Foot_6_Control_a", "SS5_Foot_6_Control_b", "SS5_Foot_6_Control_c"),
                ratio_par = 2.0) )
foot6_SS6 <- join_common( clean_common(foot6m,
                sample_names = c("SS6_Foot_6_a", "SS6_Foot_6_b"),
                control_names = c("SS6_Foot_6_Control_a", "SS6_Foot_6_Control_b"),
                ratio_par = 2.0) )
```

```{r foot7}
load("./dat/foot7.rda")
foot7m$Sample <- str_replace_all(foot7m$Sample, " ", "_")

foot7_SS5 <- join_common( clean_common(foot7m,
                sample_names = c("SS5_Foot_7_a", "SS5_Foot_7_b", "SS5_Foot_7_c"),
                control_names = c("SS5_Foot_7_Control_a", "SS5_Foot_7_Control_b", "SS5_Foot_7_Control_c"),
                ratio_par = 2.0) )
foot7_SS6 <- join_common( clean_common(foot7m,
                sample_names = c("SS6_Foot_7_a", "SS6_Foot_7_b", "SS6_Foot_7_c"),
                control_names = c("SS6_Foot_7_Control_a", "SS6_Foot_7_Control_b"),
                ratio_par = 2.0) )
```

```{r foot8}
load("./dat/foot8.rda")
foot8m$Sample <- str_replace_all(foot8m$Sample, " ", "_")

foot8_SS4 <- join_common( clean_common(foot8m,
                sample_names = c("SS4_Foot_8_a", "SS4_Foot_8_b"),
                control_names = c("SS4_Foot_8_Control_a", "SS4_Foot_8_Control_b", "SS4_Foot_8_Control_c"),
                ratio_par = 2.0) )
foot8_SS5 <- join_common( clean_common(foot8m,
                sample_names = c("SS5_Foot_8_a", "SS5_Foot_8_b", "SS5_Foot_8_c"),
                control_names = c("SS5_Foot_8_Control_a", "SS5_Foot_8_Control_b", "SS5_Foot_8_Control_c"),
                ratio_par = 2.0) )
foot8_SS6 <- join_common( clean_common(foot8m,
                sample_names = c("SS6_Foot_8_a", "SS6_Foot_8_b"),
                control_names = c("SS6_Foot_8_Control_a", "SS6_Foot_8_Control_b"),
                ratio_par = 2.0) )
```

```{r foot9}
load("./dat/foot9.rda")
foot9m$Sample <- str_replace_all(foot9m$Sample, " ", "_")

foot9_SS4 <- join_common( clean_common(foot9m,
                sample_names = c("SS4_Foot_9_a", "SS4_Foot_9_b", "SS4_Foot_9_c"),
                control_names = c("SS4_Foot_9_Control_a", "SS4_Foot_9_Control_b", "SS4_Foot_9_Control_c"),
                ratio_par = 2.0) )
foot9_SS5 <- join_common( clean_common(foot9m,
                sample_names = c("SS5_Foot_9_a", "SS5_Foot_9_b", "SS5_Foot_9_c"),
                control_names = c("SS5_Foot_9_Control_a", "SS5_Foot_9_Control_b", "SS5_Foot_9_Control_c"),
                ratio_par = 2.0) )
foot9_SS6 <- join_common( clean_common(foot9m,
                sample_names = c("SS6_Foot_9_a", "SS6_Foot_9_b", "SS6_Foot_9_c"),
                control_names = c("SS6_Foot_9_Control_a", "SS6_Foot_9_Control_b"),
                ratio_par = 2.0) )
```

## Write Excel Sheets for Rushali to Look At

```{r foot1w}
library("xlsx")
write.xlsx(x = foot1_SS2,
           file = "./dat_comp/foot1_compounds.xlsx",
           sheetName = "SS2",
           col.names = TRUE,
           row.names = TRUE,
           append = FALSE)
for(j in 3:6) {
  write.xlsx(x = get(paste0("foot1_SS", j)),
             file = "./dat_comp/foot1_compounds.xlsx",
             sheetName = paste0("SS", j),
             col.names = TRUE,
             row.names = TRUE,
             append = TRUE) 
}
```

```{r foot2w}
library("xlsx")
write.xlsx(x = foot2_SS2,
           file = "./dat_comp/foot2_compounds.xlsx",
           sheetName = "SS2",
           col.names = TRUE,
           row.names = TRUE,
           append = FALSE)
for(j in 3:6) {
  write.xlsx(x = get(paste0("foot2_SS", j)),
             file = "./dat_comp/foot2_compounds.xlsx",
             sheetName = paste0("SS", j),
             col.names = TRUE,
             row.names = TRUE,
             append = TRUE) 
}
```

```{r foot3w}
library("xlsx")
write.xlsx(x = foot3_SS5,
           file = "./dat_comp/foot3_compounds.xlsx",
           sheetName = "SS5",
           col.names = TRUE,
           row.names = TRUE,
           append = FALSE)
write.xlsx(x = foot3_SS6,
           file = "./dat_comp/foot3_compounds.xlsx",
           sheetName = "SS6",
           col.names = TRUE,
           row.names = TRUE,
           append = TRUE) 
```

```{r foot4w}
library("xlsx")
write.xlsx(x = foot4_SS5,
           file = "./dat_comp/foot4_compounds.xlsx",
           sheetName = "SS5",
           col.names = TRUE,
           row.names = TRUE,
           append = FALSE)
write.xlsx(x = foot4_SS6,
           file = "./dat_comp/foot4_compounds.xlsx",
           sheetName = "SS6",
           col.names = TRUE,
           row.names = TRUE,
           append = TRUE) 
```

```{r foot5w}
library("xlsx")
write.xlsx(x = foot5_SS2,
           file = "./dat_comp/foot5_compounds.xlsx",
           sheetName = "SS2",
           col.names = TRUE,
           row.names = TRUE,
           append = FALSE)
for(j in 3:6) {
  write.xlsx(x = get(paste0("foot5_SS", j)),
             file = "./dat_comp/foot5_compounds.xlsx",
             sheetName = paste0("SS", j),
             col.names = TRUE,
             row.names = TRUE,
             append = TRUE) 
}
```

```{r foot6w}
library("xlsx")
write.xlsx(x = foot6_SS5,
           file = "./dat_comp/foot6_compounds.xlsx",
           sheetName = "SS5",
           col.names = TRUE,
           row.names = TRUE,
           append = FALSE)
for(j in 6:6) {
  write.xlsx(x = get(paste0("foot6_SS", j)),
             file = "./dat_comp/foot6_compounds.xlsx",
             sheetName = paste0("SS", j),
             col.names = TRUE,
             row.names = TRUE,
             append = TRUE) 
}
```

```{r foot7w}
library("xlsx")
write.xlsx(x = foot7_SS5,
           file = "./dat_comp/foot7_compounds.xlsx",
           sheetName = "SS5",
           col.names = TRUE,
           row.names = TRUE,
           append = FALSE)
for(j in 6:6) {
  write.xlsx(x = get(paste0("foot7_SS", j)),
             file = "./dat_comp/foot7_compounds.xlsx",
             sheetName = paste0("SS", j),
             col.names = TRUE,
             row.names = TRUE,
             append = TRUE) 
}
```

```{r foot8w}
library("xlsx")
write.xlsx(x = foot8_SS4,
           file = "./dat_comp/foot8_compounds.xlsx",
           sheetName = "SS4",
           col.names = TRUE,
           row.names = TRUE,
           append = FALSE)
for(j in 5:6) {
  write.xlsx(x = get(paste0("foot8_SS", j)),
             file = "./dat_comp/foot8_compounds.xlsx",
             sheetName = paste0("SS", j),
             col.names = TRUE,
             row.names = TRUE,
             append = TRUE) 
}
```

```{r foot9w}
library("xlsx")
write.xlsx(x = foot9_SS4,
           file = "./dat_comp/foot9_compounds.xlsx",
           sheetName = "SS4",
           col.names = TRUE,
           row.names = TRUE,
           append = FALSE)
for(j in 5:6) {
  write.xlsx(x = get(paste0("foot9_SS", j)),
             file = "./dat_comp/foot9_compounds.xlsx",
             sheetName = paste0("SS", j),
             col.names = TRUE,
             row.names = TRUE,
             append = TRUE) 
}
```



















